Increased perceptions of autonomy through choice fail to enhance motor skill retention

### 2021-10-31 docs/r1_submission/r1_manuscript.Rmd
add: directory for r1 revisions and copy over original submission

csl               : "../apa.csl"
documentclass     : "apa7"
classoption       : "man" # man for manuscript & jou for journal
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{pdflscape}
  - \pagewiselinenumbers
  - \raggedbottom
  - \leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  - \rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \hypersetup{draft} # comment out after but right now will prevent error message about some link spanning multiple pages
---

@@

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently it has been subsumed by autonomy-support. In fact, autonomy-support is considered such a robust learning variable that it is one of three key pillars in the recently proposed "OPTIMAL" theory of motor learning [@wulf2016]. @wulf2016 argued that providing learners with opportunities for choice---considered an autonomy-supportive practice manipulation---can facilitate motor performance and learning by enhancing learner's expectancies for success, and by allowing the learner to maintain their attentional focus on the task by reducing the need for self-regulatory activity [@abdollahipour2017; @chua2018; @wulf2018]. In other words, these psychological and attentional benefits, and concomitant increases in performance and learning are a by-product of exercising choice during practice. Overall, autonomy-support is seen as a means to efficient goal-action coupling [@wulf2016] and also links the "OPTIMAL" theory with Self-Determination theory [@deci2000].

Despite its prominent role within the "OPTIMAL" theory of motor learning, there are numerous reasons to doubt the importance of autonomy-support through the provision of choice during practice. First, if the benefits are the result of having opportunities for choice then learning differences should not emerge between different self-controlled groups. However, in experiments where different groups of participants have choice over their feedback schedule, such learning differences have been found, but not limited to, when this choice is made *after* rather than before a performance attempt [@chiviacowsky2005; @carter2014], when different criteria for success are provided to participants [@chiviacowsky2012], and when the absolute number of feedback choice opportunities are limited compared to unlimited at the outset of practice [@hansen2011; @patterson2011]. Second, there has been little-to-no support for the notion that practicing in a self-controlled group is perceived as more autonomy-supportive than being in a yoked group. @ste-marie2013, for example, had participants in the self-controlled and yoked groups complete the perceived choice subscale of the Intrinsic Motivation Inventory and failed to find the expected effect of higher self-reported scores during practice in the self-controlled group. Similar findings have been reported by others [@barros2019; @carter2017a; @mckay2020b]. However, @mckay2020a recently found that practicing in a self-controlled group was perceived as more autonomy-supportive than practicing the same task in a yoked group. This increased perceived autonomy-support did not, however, translate into enhanced learning compared to the yoked group. Although the available literature does not provide substantial evidence that self-controlled practice is autonomy-supportive, this does not necessarily mean such a manipulation is not autonomy-supportive [@altman1995]. For instance, this effect may in fact be quite small and require much larger sample sizes to detect than those commonly used in motor learning experiments [@lohse2016].

>>>>

documentclass     : "apa7"
classoption       : "man" # man for manuscript & jou for journal
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{pdflscape}
  - \pagewiselinenumbers
  - \raggedbottom
  - \leftheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  - \rightheader{\textbf{THIS PREPRINT HAS BEEN SUBMITTED FOR PEER-REVIEW}}
  #- \hypersetup{draft} # comment out after but right now will prevent error message about some link spanning multiple pages
---

@@

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently it has been subsumed by autonomy-support. In fact, autonomy-support is considered such a robust learning variable that it is one of three key pillars in the recently proposed "OPTIMAL" theory of motor learning [@wulf2016]. @wulf2016 argued that providing learners with opportunities for choice---considered an autonomy-supportive practice manipulation---can facilitate motor performance and learning by enhancing learner's expectancies for success, and by allowing the learner to maintain their attentional focus on the task by reducing the need for self-regulatory activity [e.g., @abdollahipour2017; @chua2018; @wulf2018]. In other words, these psychological and attentional benefits, and concomitant increases in performance and learning are a by-product of exercising choice during practice.

Despite its prominent role within the "OPTIMAL" theory of motor learning, there are reasons to doubt the importance of autonomy-support through the provision of choice during practice. First, if the benefits are primarily the result of having opportunities for choice then learning differences should not emerge between different self-controlled groups. However, in experiments where different groups of participants have choice over their feedback schedule, such learning differences have been found when this choice is made *after* rather than before a performance attempt [@chiviacowsky2005; @carter2014], when different criteria for success are provided to participants [@chiviacowsky2012], and when the absolute number of feedback choice opportunities are limited compared to unlimited at the outset of practice [@hansen2011; @patterson2011]. Second, the notion that practicing in a self-controlled group is in fact more autonomy-supportive than practicing in a yoked group has predominantly been assumed rather than supported empirically.

there has been limited support for the notion that practicing in a self-controlled group is in fact perceived as more autonomy-supportive than being in a yoked group.

@ste-marie2013, for example, had participants in the self-controlled and yoked groups complete the perceived choice subscale of the Intrinsic Motivation Inventory and failed to find the expected effect of higher self-reported scores during practice in the self-controlled group. Failures to detect increased perceptions of autonomy in self-controlled groups relative to yoked groups have also been reported by others [@barros2019; @carter2017a; @mckay2020b]. However, @mckay2020a recently found that practicing in a self-controlled group was perceived as more autonomy-supportive than practicing the same task in a yoked group. This increased perceived autonomy-support did not, however, translate into enhanced learning compared to the yoked group. Although the available literature does not provide substantial evidence that self-controlled practice is autonomy-supportive, this does not necessarily mean such a manipulation is not autonomy-supportive [@altman1995]. For instance, this effect may in fact be quite small and require much larger sample sizes to detect than those commonly used in motor learning experiments [@lohse2016].
